---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# selenium

<!-- badges: start -->
[![R-CMD-check](https://github.com/ashbythorpe/selenium-r/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/ashbythorpe/selenium-r/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

selenium is a tool for the automation of web browsers. It is a low-level interface 
to the [WebDriver](https://w3c.github.io/webdriver/) specification, and an up-to-date
alternative to [RSelenium](https://github.com/ropensci/RSelenium).

## Installation

You can install the development version of selenium from GitHub:

``` r
# install.packages("pak")
pak::pak("ashbythorpe/selenium-r")
```

However, you must also have a selenium server installed and running (see below).

## Starting the server
A selenium instance consists of two parts: the client and the server.
The selenium package *only provides the client*. This means that you
have to start the server yourself. 

To do this you must:

* Install a browser that you want to automate (e.g. Chrome, Firefox, Edge).
* Download [Java](https://www.java.com/).

There are many different ways to download and start the server, one of which
is provided by selenium:

```{r setup}
library(selenium)
```

```r
server <- selenium_server()
```

This will download the latest version of the server and start it. You can
also do this manually if you want:

1. Download the latest `.jar` file for Selenium Server. Do this by navigating to
   the latest GitHub release page (<https://github.com/SeleniumHQ/selenium/releases/latest/>),
   scrolling down to the **Assets** section, and downloading the file named 
   `selenium-server-standalone-<VERSION>.jar` (with `<VERSION>` being the latest release version).
2. Make sure you are in the same directory as the file you downloaded.
3. In the terminal, run `java -jar selenium-server-standalone-<VERSION>.jar standalone --selenium-manager true`,
   replacing `<VERSION>` with the version number that you downloaded. This will download
   any drivers you need to communicate with the server and the browser, and start the server.

There are a few other ways of starting Selenium Server:

* Using docker to start the server. See <https://github.com/SeleniumHQ/docker-selenium>.
  This is recommended in a non-interactive context (e.g. GitHub Actions).
* Using the `wdman` package to start the server from R, using `wdman::selenium()`. Note
  that at the time of writing, this package does not work with the latest version of
  Chrome.

## Starting the client
Once a server instance has started, move to R and load selenium.
```{r selenium}
library(selenium)
```

Client sessions can be started using `SeleniumSession$new()`
```r
session <- SeleniumSession$new()
```

By default, this will connect to Firefox, but you can use the `browser` argument to specify
a different browser if you like.

```{r session}
session <- SeleniumSession$new()
```

Here, we use the `capabilities` argument to specify options for the browser.
Here, the `remote-debugging-port` argument to Chrome is used to make sure the port
that the browser uses does not conflict with any others (and may be
necessary if Chrome is not working by default).

```r
session <- SeleniumSession$new(
  browser = "chrome",
  capabilities = list(
    `goog:chromeOptions` = list(
      args = list("remote-debugging-port=9222")
    )
  )
)
```

If this doesn't work, please see the
[Debugging Selenium](https://ashbythorpe.github.io/selenium-r/articles/debugging.html)
article for more information.

<!-- TODO: Make this vignette --->

## Usage

Once the session has been successfully started, you can use the session
object to control the browser. Here, we dynamically navigate through
the R project homepage.

```{r example}
session$navigate("https://www.r-project.org/")
session$
  find_element(using = "css selector", value = ".row")$
  find_element(using = "css selector", value = "ul")$
  find_element(using = "css selector", value = "a")$
  click()

session$
  find_element(using = "css selector", value = ".row")$
  find_elements(using = "css selector", value = "div")[[2]]$
  find_element(using = "css selector", value = "p")$
  get_text()

session$close()
```

For a more detailed introduction to using selenium, see the
[Getting Started](https://ashbythorpe.github.io/selenium-r/articles/selenium.html)
article.
<!-- TODO: Make this vignette --->

Note that selenium is low-level and mainly aimed towards developers. If you are
wanting to use browser automation for web scraping or testing, you may want to
take a look at [selenider](https://github.com/ashbythorpe/selenider) instead.
