---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# selenium

<!-- badges: start -->
[![R-CMD-check](https://github.com/ashbythorpe/selenium-r/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/ashbythorpe/selenium-r/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/ashbythorpe/selenium-r/branch/main/graph/badge.svg)](https://app.codecov.io/gh/ashbythorpe/selenium-r?branch=main)
<!-- badges: end -->

selenium is a tool for the automation of web browsers. It is a low-level interface 
to the [WebDriver](https://w3c.github.io/webdriver/) specification, and an up-to-date
alternative to [RSelenium](https://github.com/ropensci/RSelenium).

## Installation

You can install the development version of selenium from GitHub:

``` r
pak::pak("ashbythorpe/selenium")
```

However, you must also have a selenium server installed and running (see below).

## Usage
A selenium instance consists of two parts: the client and the server.
The selenium package *only provides the client*. This means that you
have to start the server yourself. 

One of the ways to do this is the following:

1. Install a browser that you want to automate (e.g. Chrome, Firefox, Edge).
2. Download [Java](https://www.java.com/).
3. Download the latest `.jar` file for Selenium Server. Do this by navigating to
   the latest GitHub release page (<https://github.com/SeleniumHQ/selenium/releases/latest/>),
   scrolling down to the **Assets** section, and downloading the file named 
   `selenium-server-standalone-<VERSION>.jar` (with `<VERSION>` being the latest release version).
4. Make sure you are in the same directory as the file you downloaded.
5. In the terminal, run `java -jar selenium-server-standalone-<VERSION>.jar standalone --selenium-manager true`,
   replacing `<VERSION>` with the version number that you downloaded. This will download
   any drivers you need to communicate with the server and the browser, and start the server.

There are a few other ways of starting Selenium Server:

* Using docker to start the server. See <https://github.com/SeleniumHQ/docker-selenium>.
  This is recommended in a non-interactive context (e.g. GitHub Actions).
* Using the `wdman` package to start the server from R, using `wdman::selenium()`. Note
  that at the time of writing, this package does not work with the latest version of
  Chrome.

Once a server instance has started, move to R and load selenium.
```{r selenium}
library(selenium)
```

Client sessions can be started using `SeleniumSession$new()`
```r
session <- SeleniumSession$new()
```

By default, this will connect to Chrome, but you can use the `browser` argument to specify
a different browser if you like.

```{r session}
session <- SeleniumSession$new(browser = "firefox")
```

If this doesn't work, you may have to specify the `port` and `host` arguments manually.
If you used the first method above to start the server, you should have seen a message
like this output in the terminal:
`INFO [Standalone.execute] - Started Selenium Standalone ... (revision ...): http://<IP>:<PORT>`
The URL at the end of this message can be used to extract an IP address and a port number, which
can then be passed into the `host` and `port` arguments. For example, if the URL was:
`http://172.17.0.1/4444`, you would run:
```r
session <- SeleniumSession$new(host = "172.17.0.1", port = 4444)
```

Once the session has been successfully started, you can use the session
object to control the browser.

```{r example}
session$navigate("https://www.r-project.org/")
session$
  find_element(using = "css selector", value = ".row")$
  find_element(using = "css selector", value = "ul")$
  find_element(using = "css selector", value = "a")$
  click()

session$
  find_element(using = "css selector", value = ".row")$
  find_elements(using = "css selector", value = "div")[[2]]$
  find_element(using = "css selector", value = "p")$
  get_text()

session$close()
```

For a more detailed introduction to using selenium, see the Getting
Started vignette (`vignette("selenium")`).
<!-- TODO: Make this vignette --->

Note that selenium is low-level and mainly aimed towards developers. If you are
wanting to use browser automation for web scraping or testing, you may want to
take a look at [selenider](https://github.com/ashbythorpe/selenider) instead.
