---
title: "Debugging selenium"
---

```{r, include = FALSE}
available <- selenium::selenium_server_available()
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = available
)
```

```{r eval = !available, echo = FALSE, comment = NA}
if (!available) {
  message("Selenium server is not available.")
}
```

Web automation is a complex and fragile task, with many ways to go wrong.
This article describes some common errors and pitfalls, and how you might
go about resolving them.

## Starting the client

If initializing a `SeleniumSession` fails, it is often useful to look at
the logs from the server. If you ran the `java -jar ...` command manually, then
you should be able to see the logs, but if you used `selenium_server()`, then
you need to use the `stdout` argument to capture the output of the command:

```r
# Where `log.txt` is an already created file where you want to store the logs.
selenium_server(stdout = "log.txt")
```

You can then look at the log file to see the output from the command.

### Port and IP address
One reason why you can't connect to the server is that the port and IP address you
are using is wrong.

To get the IP address and port number that the server is using, you need to look
at the server logs. You should see a line like:
`INFO [Standalone.execute] - Started Selenium Standalone ... (revision ...): http://<IP>:<PORT>`

The URL at the end of this message can be used to extract an IP address and a port number, which
can then be passed into the `host` and `port` arguments. For example, if the URL was:
`http://172.17.0.1/4444`, you would run:
```r
session <- SeleniumSession$new(host = "172.17.0.1", port = 4444)
```

### Using a different debugging port for Chrome
If you are using Chrome, and you see a browser open, but the call to
`SeleniumSession$new()` times out, you may need to use a different debugging port.
For example:

```r
session <- SeleniumSession$new(
  browser = "chrome",
  capabilities = list(
    `goog:chromeOptions` = list(
      args = list("remote-debugging-port=9222")
    )
  )
)
```

### Increasing /dev/shm/ size when using docker
If you are running selenium using docker, you may need to increase the size of
`/dev/shm/` to avoid running out of memory. This issue usually happens when
using Chrome, and usually results in a message like
`session deleted because of page crash`.

You can use the `--shm-size` to the selenium docker images to fix this issue.
For example:
`docker run --shm-size="2g" selenium/standalone-chrome:<version>`

## Other common errors

### Stale element reference errors
At some point, when using selenium, you will encounter the following error:

```r
#> Error in `element$click()`:
#> ! Stale element reference.
#> ✖ The element with the reference <...> is not known in the current browsing context
#> Caused by error in `httr2::req_perform()`:
#> ! HTTP 404 Not Found.
#> Run `rlang::last_trace()` to see where the error occurred.
```
This error is common when automating a website. Selenium is telling you that
an element which you previously identified no longer exists. In all websites,
especially complex ones, the DOM will be constantly updating itself, constantly
invalidating references to elements. This error is a particularly annoying one,
as it can happen at any time and is impossible to predict.

One way to deal with this error is to use elements as soon as they are created,
only keeping references to elements if you are sure that they will not be
invalidated. For example, if you want to click the same element twice, with
a second-long gap in between, you may want to consider fetching the element
once for each time, rather than sharing the reference between the actions.

However, this solution is not infallible. If you find yourself encountering
this error a lot, it may be a sign that a more high-level package, that can
deal with this issue (e.g. [selenider](https://github.com/ashbythorpe/selenider)),
is needed.
