---
title: "Getting Started with selenium"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

selenium is low-level tool, so can be complex to use. This vignette gives an
overview of the important functions of selenium, and how to use them.

```{r setup}
library(selenium)
```

First, let's create a new session (assuming that you have a server running):
```{r session}
session <- SeleniumSession$new()
```

First, navigate to the website you want to automate with `SeleniumSession$navigate()`:

```{r navigate}
session$navigate("http://www.tidyverse.org")
```

## Navigation

Use `SeleniumSession$back()` and `SeleniumSession$forward()` to navigate
back and forward in the navigation history, and
`SeleniumSession$refresh()` to reload the current page

```{r history}
session$back()

session$forward()

session$refresh()

session$current_url()
```

## Elements

To interact with a webpage, you first need to find the elements you want to
interact with. You can do this using `SeleniumSession$find_element()` and
`SeleniumSession$find_elements()`. Both functions take two arguments:
`using` and `value`. `using` specifies the method you want to use to find the
element, while `value` is the string you want to search for.

By default, CSS selectors are used (`value = "css selector"`), as they provide
a simple and concise way to find elements:

* Use `"name"` to select an element with the tag name `name`. (e.g. `div`).
* Use `"#id"` to select an element with the id `id`.
* Use `".class"` to select an element with the class `class`.
* Use square brackets to select elements using other properties.
  For example, `"input[type='text']"` selects a text input element.

For more information about using CSS selectors, see:
<https://developer.mozilla.org/en-US/docs/Web/CSS/CSS_selectors/Selectors_and_combinators>.

Another notable way to select elements is using XPaths. Xpaths can be slightly more
verbose than CSS selectors, but are more powerful. Use `using = "xpath"` to use
XPaths.

* Use `//tagname` to select an element with the tag name `tagname`.
* Use `//tagname[@attribute='value']` to select an element with the tag name
  `tagname` and the attribute `attribute` with the value `value`.
* Use `//tagname/childtagname` (single slash) to select an element that is a direct
  child of the `tagname` element. Similarly, `/tagname` can be used to select
  an element that is at the top level of the DOM.
* Use `//tagname//childtagname` to select an element that is a child of the 
  `tagname` element, but does not have to be directly below it in the DOM.

To learn more about how to use Xpaths, see:
<https://www.w3schools.com/xml/xpath_intro.asp>

The other values that `using` can take are fairly self-explanatory (e.g. `"tag name"`)
will match elements with the tag name given by `value`.

Once you have an element, you can locate further sub-elements using
`WebElement$find_element()` and `WebElement$find_elements()`, which will
now select from the element's children.
This is especially useful when a website does not make much use of
classes and ids to label its elements uniquely.

For example, the following code will find all the hex images on the Tidyverse
home page:

```{r hex}
hex_stickers <- session$
  find_element(using = "css selector", value = ".hexBadges")$
  find_elements(using = "css selector", value = "img")
```

## Element properties
Once you have an element, there are many properties that you can request from
it.

```{r get_element}
element <- hex_stickers[[1]]
```

Use `WebElement$get_tag_name()` to get the tag name of an element:

```{r tag_name}
element$get_tag_name()
```

Use `WebElement$get_text()` to get the text inside of an element:

```{r text}
element$get_text()
```

`WebElement$get_attribute()` and `WebElement$get_property()` will return
the value or property of an element (see the documentation for the
difference between the two).

```{r get_attribute}
element$get_attribute("src")

element$get_property("src")
```

`WebElement$get_css_property()` will return the computed CSS property of an
element:

```{r get_css_property}
element$get_css_property("background-color")
```

Use `WebElement$get_rect()` to get the coordinates and size of an element:

```{r rect}
element$get_rect()
```

Use `WebElement$is_selected()`, `WebElement$is_enabled()` and
`WebElement$is_displayed()` to check if an element is currently selected,
enabled or visible.

```{r check}
element$is_selected()

element$is_enabled()

element$is_displayed()
```

## Interacting with an element
There are many different ways to interact with an element.

Use `WebElement$click()` to click on an element:

```{r click}
element$click()
```

Use `WebElement$send_keys()` to send text to an input element:

```{r send_keys}
input <- session$find_element(using = "css selector", value = "input[type='search']")

input$send_keys("filter")
```

You can use the `keys` object to send more advanced key combinations to an element.

For example, the below code selects all the text inside the element (`<Ctrl-A>`), and
then deletes it.

```{r keys}
input$send_keys(keys$ctrl, "a", keys$backspace)
```

However, generally, if you want to clear an input element, use `WebElement$clear()`.

```{r clear}
input$clear()
```

## Custom JavaScript
Use `SeleniumSession$execute_script()` to execute custom JavaScript. Any extra arguments
will be passed into the script through the `arguments` array, and items can be returned
explicitly using `return`. `WebElements` can also be passed in and returned.

```{r js}
session$execute_script("return arguments[0] + arguments[1];", 2, 2)
```

```{r js_element}
session$execute_script("return arguments[0].parentElement;", element)
```

`SeleniumSession$execute_async_script()` is provided for when you need to interact
with an asynchronous JavaScript API, but is much less commonly used.

## Windows and Tabs
A `SeleniumSession` object is not limited to a single window/tab. Use
`SeleniumSession$new_window()` to create a new window or tab, using the
`type` argument to choose between the two:

```{r new_window}
tab <- session$new_window(type = "tab")

window <- session$new_window(type = "window")

tab

window
```

Each window/tab has its own handle: a string that uniquely identifies it.
Use `SeleniumSession$window_handle()` to get the handle of the current
window, and `SeleniumSession$window_handles()` to get the handle of
every open window/tab.

```{r window_handle}
current_handle <- session$window_handle()

current_handle

session$window_handles()
```

When a new window/tab is created, it is not automatically switched to. Use
`SeleniumSession$switch_to_window()` to switch to a window/tab with a specific
handle:

```{r switch_to_window}
session$switch_to_window(window$handle)
```

Use `SeleniumSession$close_window()` to close a window/tab:

```{r close_window}
session$close_window()

# Switch to an open window
session$switch_to_window(current_handle)
```

## iframes and the Shadow DOM
There are many methods that web developers use to embed self-contained
HTML documents or elements into a page.

One of these is an iframe element, allowing a nested browsing context to
be embedded into a page. For example, iframes are often used to embed
video players into a site. Use `SeleniumSession$switch_to_frame()` to switch
to an frame and select elements within it. You can pass in the index of the
frame, or the `iframe` element itself.

```{r switch_to_frame}
session$navigate("https://www.youtube.com")

iframe <- session$find_element(using = "css selector", value = "iframe")

session$switch_to_frame(iframe)
```

Use `SeleniumSession$switch_to_parent_frame()` to switch to the parent
frame of the current one, or use `SeleniumSession$switch_to_frame()`
with no arguments to switch to the top-level browsing context:

```{r switch_to_parent_frame}
session$switch_to_parent_frame()
```

The Shadow DOM is another way for websites to embed content into a page,
this time into any element. If an element is a shadow root, use
`WebElement$shadow_root()` to get the corresponding `ShadowRoot`
object. You can then find sub-elements within the shadow root
using `ShadowRoot$find_element()` and `ShadowRoot$find_elements()`.

## Closing the session
Always remember to close the session after you have finished using it:

```{r close}
session$close()
```
